\documentclass{article}

\usepackage{amsmath}
\usepackage{amsfonts}

\usepackage[margin=1in]{geometry}

\begin{document}

\noindent {\Large \textbf{PROBLEM 1.2}}  (Guillot, 2018, p. 21) \\\\
\indent Let $F$ be a field of characteristic $p \neq 2$. Let $a \in F^\times{\setminus}F^{\times2}$, and $K = F[\sqrt{a}]$. \\\\
\indent We do not assume that $\sqrt{-1} \in{F}$. \\\\
\\
\textbf{PART 1} \\\\
\indent Suppose there exists an extension $L/F$ with $K \subset L$. \\\\
\indent Suppose further that $L/F$ is Galois with $\textbf{Gal}(L/F) \cong C_4$, the cyclic group of order $4$. \\\\
\indent Show that there exists an $\alpha \in K$ such that $N_{K/F}(\alpha) = -1$. \\\\
\indent \emph{Hint}: try $\alpha = \dfrac{\theta - \sigma^2(\theta)}{\sigma(\theta) - \sigma^3(\theta)}$. \\\\
\\
\textbf{SOLUTION 1.2.1} \\\\
\noindent Since $\textbf{Gal}(L/F)$ is cyclic of degree $4$, we have: \\\\
\indent $\bullet$ \indent $L = F[\sqrt[4]{b}]$ for some $b \in{F^{\times}}$, and \\\\
\indent $\bullet$ \indent the order of $[b]$ in ${F^\times/F^{\times4}}$ is $4$. \\\\
Here $[b]$ is the image of $b \in{F^{\times}}$ via the group homomorphism $F^{\times} \rightarrow F^{\times}/F^{\times4}$. \\\\
This is an application of \textbf{Proposition 1.7}, part (2) in (Guillot, 2018, p. 5). \\\\
Already we can deduce that $L$ contains a primitive $4^{th}$ root of unity: \\\\
\indent By Galois theory, \emph{any} irreducible polynomial $f(x)$ over F with one root in $L$ has \emph{all} its roots in $L$. \\\\
\indent \indent By ``all" roots, we of course mean every root of $f$ in the algebraic closure $\overline{F}$. \\\\
\indent \indent See \textbf{Proposition 3.28}, part (4) in (Morandi, 1996, p.36) for details. \\\\
\indent We are free to let $b = a$ and $f(x) = x^4 - a$. The latter is irreducible over $F$: \\\\ 
\indent \indent $\bullet$ \indent $F$ is not a field of characteristic dividing the degree of $f$. \\\\
\indent \indent $\bullet$ \indent The problem states that $a$ is not a square in $F^{\times}$. \\\\
\indent \indent $\bullet$ \indent Since $a \notin F^{\times2}$, nor is it a $4^{th}$ power in $F^{\times}$.\\\\ 
\indent \indent All this shows that there is no linear or quadratic factor that can be split off from $f$.\\\\
\indent \indent Such must be the case for a reducible polynomial of degree $4$. \\\\
\indent So $\sqrt[4]{a}$ is a root of $f$ lying in $L = F[\sqrt[4]{a}]$. By \textbf{Proposition 3.28} all other roots of $f$ lie in $L$. \\\\
\indent That would include $i\sqrt[4]{a}$, and $i = \sqrt{-1}$ is a primitive $4^{th}$ root of unity as promised. \\\\
We further investigate this field extension, as we often do, by: \\\\
\indent $\bullet$ \indent analyzing the (mercifully linear) action of $\textbf{Gal}(L/F)$ on $L$, and \\\\
\indent $\bullet$ \indent using norms to construct field elements (or, to prove that they don't exist). \\\\
Taking Guillot's hint: let $\alpha$ be of the form $\dfrac{\theta - \sigma^2(\theta)}{\sigma(\theta) - \sigma^3(\theta)}$. \\\\
We will take $\theta$ to be a primitive $4^{th}$ root of unity, and $\sigma$ to be a generator of $\textbf{Gal}(L/F) \cong C_4$. \\\\
Clearly, $\alpha$ lies in $L$. Since $\sigma$ acts linearly on $L$, we have: \\\\
\indent $\sigma(\alpha) = \dfrac{\sigma(\theta) - \sigma^3(\theta)}{\sigma^2(\theta) - \theta} = -1/\alpha$. $(\ast)$ \\\\
Note that the norm of $\alpha$ with respect to $L/F$ is given by $N_{L/F}(\alpha) = \alpha \, \sigma(\alpha) \, \sigma^2(\alpha)\, \sigma^3(\alpha)$. \\\\
On the other hand, $N_{L/K}(\alpha)$ involves only those powers of $\sigma$ that lie in $\textbf{Gal}(L/K)\triangleleft{\textbf{Gal}(L/F)}$. \\\\
\indent Since $[L:F] = [L:K][K:F] = 4$, $[L:K]$ must be $2$. \\\\
\indent So $\textbf{Gal}(L/K)$ is a subgroup of degree $2$ in $G = \textbf{Gal}(L/F).$ \\\\
\indent The only such subgroup is given by $\{1_G, \sigma^2\}$. \\\\
So $N_{L/K}(\alpha) = \alpha \, \sigma^2(\alpha)$, and this is equal to $\alpha^2$ by $(\ast)$ above. \\\\
Since the image of $N_{L/K}$ is $K$, we now know that $\alpha^2 \in K$. \\\\
This implies that $\alpha \in K$. Now, we ask for $N_{K/F}(\alpha)$ once we know $\textbf{Gal}(K/F)$. \\\\
\indent The problem states that $K$ is generated as $F[\sqrt{a}]$ for some $a \notin F^{\times2}$ \\\\
\indent So every element of $K$ is of the form $x + \sqrt{a}y$ for some $x, y \in F$ and $[K:F] = 2$. \\\\
\indent Any extension $K/F$ of degree $2$ over a field of characteristic $p \neq 2$ is Galois. \\\\ 
\indent Also, $\textbf{Gal}(K/F) \cong C_2$, and the non-identity  --- conjugation --- element sends $\sqrt{a}$ to $-\sqrt{a}$. \\\\
Notice that conjugation is the same as the restriction of $\sigma$ to $K$, denoted by $\sigma|_K$. \\\\
\indent In other words, $\sigma$ is the preimage of conjugation in $K$ via $\dfrac{\textbf{Gal}(L/F)}{\textbf{Gal}(L/K)} \cong \textbf{Gal}(K/F)$. \\\\
\indent To see this more clearly, observe that $\sigma$ cyclically permutes the roots of $f(x)$ above. \\\\
\indent \indent We have $\sigma(\sqrt[4]a) = \theta \sqrt[4]{a}$. Therefore $\sigma(\sqrt[4]{a}) \, \sigma(\sqrt[4]{a}) = - \sqrt[4]{a} \, \sqrt[4]{a} = - \sqrt{a}$. \\\\
\indent \indent By the linearity of $\sigma$, the leftmost side reduces so that $\sigma(\sqrt{a}) = - \sqrt{a}$.\\\\
Finally, the norm of $\alpha$ with respect to $K/F$ is $\alpha \, \sigma|_K(\alpha)$. \\\\
By $(\ast)$ above, $\alpha \, \sigma|_K(\alpha) = N_{K/F}(\alpha) = -1$. $\square$ \\\\

\noindent \textbf{PART 2} \\\\
\indent Let $\alpha \in K$ be an element with norm $N_{K/F}(\alpha) = -1$. \\\\
\indent Show that this implies the existence of a cyclic extension $L/F$ of degree 4. \\\\
\indent \emph{Hint}: The case when $\alpha \in F$ must be treated separately. When $\alpha \notin F$, try $L = K[\sqrt{1+\alpha^2}]$. \\\\
\\
\textbf{SOLUTION 1.2.2} \\\\
\noindent The element $\alpha$ is of the form $x + \sqrt{a} y$ for some $x, y \in F$. First, suppose $\alpha \in F$: \\\\
\indent We have $y = 0$. Since $N_{K/F}(\alpha) = -1$ by assumption, we also have $x^2 = -1$. \\\\
\indent This implies that $i = \sqrt{-1} \in F$. So $F$ contains a primitive $4^{th}$ root of unity: \\\\
\indent Therefore we may apply a (limited) version of $\emph{The Fundamental Theorem of Kummer Theory}$: \\\\ 
\indent \indent Subgroups of order $n \geq 1$ in $F^\times/F^{\times n}$ give rise to Galois extensions of degree $n$. \\\\
\indent \indent Let $[b]$ be such a subgroup. Let $L/F$ be such an extension. \\\\
\indent \indent Then $L$ is generated as $F[\sqrt[n]{b}]$ for any $b$ in $[b]$. \\\\
\indent \indent See \textbf{Theorem 1.25} in (Guillot, 2018, p. 14) for the full theorem. \\\\
\indent It suffices to show that such a subgroup and a generating element exist for $n = 4$. \\\\
\indent Before we prove this, observe that the exponent of $F^\times/F^{\times4}$ is $4$. $(\ast)$ \\\\
\indent \indent Every element of $[b]$ is of the form $bx^{4}$ for some $b \in F^\times{\setminus}F^{\times4}$ and $x \in F^\times$. \\\\
\indent \indent Furthermore, $bx^4$ is written in lowest terms, so that $b$ does not contain a $4^{th}$ power. \\\\
\indent \indent Let $m$ be the order of $[b]$ in $F^\times/F^{\times4}$.  If $m > 4$,  there exists some $c = (bx^{4})^m \in F^{\times 4}$. \\\\
\indent \indent This implies that $b^m = b^{m-4}b^4 = 1$, contradicting that $c$ is given in lowest terms. \\\\
\indent By group theory, the order of any subgroup $[b]$ in $F^\times/F^{\times4}$ must divide the exponent $4$. \\\\
\indent Now it is clear that we may choose $b = a$ and $[b] = [a]$: \\\\
\indent \indent Since $a$ is not a square or a $4^{th}$ power in $F^\times$, the order of $[a]$ cannot be $2$ or $1$ (resp.). \\\\
\indent \indent So the order of $[a]$ is $4$, and a cyclic extension $L/K$ of degree $4$ exists. \\\\
Now suppose $\alpha \notin F$. We take Guillot's hint and try $L = K[\sqrt{1+\alpha^2}]$. \\\\
\indent Note that any extension of degree $2$ over a field of characteristic $p \neq 2$ is Galois. \\\\
\indent Since $a \notin F^{\times2}$ and $K = F[\sqrt{a}]$, $K/F$ is Galois with $\textbf{Gal}(K/F) \cong C_2$. \\\\
\indent Let $\sigma$ be the non-identity. Since $\sigma$ acts linearly, we have: \\\\
\indent \indent $N_{K/F}(1 + \alpha^2) = (1 + \alpha^2) \, \sigma(1 + \alpha^2) = (1 + \alpha^2) \, (1 + \sigma(\alpha)^2).$ \\\\
\indent Recall that $N_{K/F}(\alpha) = \alpha \, \sigma(\alpha) = -1$, and so $\alpha^2 \, \sigma(\alpha)^2 = 1$. \\\\
\indent Applying these formulae, we get $N_{K/F}(1 + \alpha^2) = (\alpha - \sigma(\alpha))^2$. \\\\
\indent If $1 + \alpha^2$ itself lies in $K^{\times2}$, then $N_{K/F}(1 + \alpha^2) = N_{K/F}(\beta)^2$ for some $\beta \in K$. \\\\
\indent But $N_{K/F}(\beta) \in F$, which implies that $N_{K/F}(1 + \alpha^2)$ also lies in $F^{\times2}$. \\\\
\indent This cannot be the case since $(\alpha - \sigma(\alpha))^2 = 4y^2a$  does not lie in $F^{\times2}$. \\\\
\indent \indent While $4y^2$ is a square, recall that we have assumed $a \notin F^{\times2}$. \\\\
\indent Therefore, $1 + \alpha^2$ is not a $4^{th}$ power in $K^{\times}$. Nor is it a square. \\\\
\indent Using the same argument at $(\ast)$ above: \\\\
\indent \indent The exponent of $K^\times/K^{\times4}$ is $4$. Let $[b]$ be a subgroup. \\\\
\indent \indent The order of $[b]$ in $K^\times/K^{\times4}$ divides the exponent $4$. \\\\
\indent Since the order of $[1 + \alpha^2]$ in $K^\times/K^{\times4}$ is not $1$ or $2$, it must be $4$. \\\\
\indent Consider the element $1 + x^2 \in K^\times$ where $x$ is the component of $\alpha$ lying in $F^\times$ (i.e., $\alpha = x + \sqrt{a} y$). \\\\
\indent This is also not a square or $4^{th}$ power in $K^\times$. \\\\
\indent \indent To see this, observe that the component of $w^2 = (u + \sqrt{a} v)^2$ lying in $F$ is $(u^2 + ay^2)$. \\\\
\indent \indent But if $w$ lies in $F$, $(u^2 + ay^2) = u^2$ with $y = 0$. \\\\
\indent So the subgroup $[1 + x^2]$ has order $4$ in $K^\times/K^{\times4}$. The upshot is that $1 + x^2$ lies in $F^\times$: \\\\
\indent We may now consider the preimage of $[1 + x^2]$ via the \emph{injective} homomorphism $g: F^\times/F^{\times4} \rightarrow K^\times/K^{\times4}$. \\\\ 
\indent By group theory, the order of $g^{-1}([1 + x^2])$ in $F^\times/F^{\times4}$ is equal to the order of $[1 + x^2]$ in $K^\times/K^{\times4}$. \\\\
\indent So we have constructed a subgroup of order $4$ in $F^\times/F^{\times4}$, namely $g^{-1}([1 + x^2]) = [1 + x^2]$. \\\\
\indent By the fundamental theorem, any $4^{th}$ root of an element in $[1 + x^2]$ generates $L/F$. $\square$ \\\\
\\

\noindent \textbf{PART 3} \\\\
\indent Show that the following are equivalent: \\\\
\indent \indent (a) \indent There exists an $\alpha \in K$ with $N_{K/F}(\alpha) = -1$, and \\\\
\indent \indent (b) \indent $a$ is a sum of two squares in $F$ \\\\
\\
\textbf{SOLUTION 1.2.3} \\\\
\noindent Again, note that any extension of degree $2$ over a field of characteristic $p \neq 2$ is Galois. \\\\
\indent Since $a \notin F^{\times2}$ and $K = F[\sqrt{a}]$, $K/F$ is Galois with $\textbf{Gal}(K/F) \cong C_2$. \\\\
\indent Let $\sigma$ be the non-identity, which sends $\sqrt{a}$ to $\sqrt{-a}$. \\\\
If $a = x^2 + y^2$ for some $x, y \in F$, then $\dfrac{x^2-a}{y^2} = -1$. We also have \\\\
\indent \indent $-1 =\dfrac{x + \sqrt{a}}{y} \cdot \dfrac{x - \sqrt{a}}{y} =\dfrac{x + \sqrt{a}}{y} \cdot \sigma(\dfrac{x + \sqrt{a}}{y})$. \\\\
Suppose $\alpha = \dfrac{x + \sqrt{a}}{y} \in K$. Then the rightmost side of the equation above is just $N_{K/F}(\alpha)$. \\\\
This proves that (b) $\implies$ (a). \\\\
Conversely, if $N_{K/F}(\alpha) = -1$, then $(x + \sqrt{a}y) \cdot (x - \sqrt{a}y) = x^2 - ay^2 = -1$. \\\\
We can rearrange this as $a = \left(\dfrac{1}{y}\right)^2 + \left(\dfrac{x}{y}\right)^2$, which is a sum of squares in $F$. \\\\
This proves that (a) $\implies$ (b). $\square$ \\\\
\pagebreak

\noindent {\Large \textbf{REFERENCES}} \\\\
\noindent Guillot, Pierre. (2018). \emph{A Gentle Course in Local Class Field Theory: Local Number Fields, Brauer Groups, Galois Cohomology}. Cambridge: Cambridge University Press. \\\\
\noindent Morandi, Patrick. (1996). \emph{Field and Galois Theory}. Graduate Texts in Mathematics, vol 167. Springer, New York, NY. \\\\

\end{document}